The CrypTool Script: \\ Cryptography, Mathematics and more \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free eLearning program CrypTool\\ (8th edition -- distributed with CrypTool version 1.4.10)} \vspace{60pt}

### 2007-07-25 script/en/chapters/paper_and_pencil.tex
crypto2020 with bibtext added; German Umlaute enabled; some tables narrowed
CT-scripts: Minor corrections -> should be final now for 1.4.10

\item {\bf Shift cipher, for example Caesar cipher}\footnote{In CrypTool
   this method  can be find at three different places in the menu tree:\\
   - {\bf Crypt \textbackslash{} Symmetric (classic) \textbackslash{} 
      Caesar / ROT13}\\
   - {\bf Analysis \textbackslash{} Symmetric Encryption (classic)
     \textbackslash{} Ciphertext only \textbackslash{} Caesar} \\
   - {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     using ANIMAL \textbackslash{} Caesar}. } 
   \cite{Singh2001}\index{Caesar}%
   : Plaintext alphabet and ciphertext alphabet are shifted against each 
   other by a determined number of letters. Using the Caesar cipher means 
   shifting letters about three positions.

@@

\item {\bf Nihilist Substitution}\footnote{An animation of this Nihilist method
   can be find in CrypTool at the menu item
     {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     using ANIMAL \textbackslash{} Nihilist}. }
   \cite{ACA2002}\index{Nihilist substitution}: 
   Insert the alphabet into a 5x5-matrix and replace 
   every letter of the message with the two corresponding digits. The 
   resulting numbers are written into a grille. For this purpose, a 
   keyword is chosen and placed above the columns of the grille. Its 
   letters are substituted by numbers, too. The ciphertext results from 
   adding the numbers of the plaintext and the numbers of the keyword.
   Numbers between 100 and 110 are transformed to numbers between 00 and
   10, so that each letter is represented by a two-digit number.

@@

   It is ostentatious, how often the numbers 1 and 1 appear,
   but this will be fixed with the following version.\\

@@

   \begin{center}
   \begin{tabular}{|ccc|ccc|ccc|ccc|ccc|ccc|}
   \hline		
	co & mb & in & in & gs & ub & st & it & ut & io & na & nd & tr & an & sp & os & it & io\\
   \hline
	3  & 4  & 9  & 9  & 7  & 3  & 1  & 9  & 3  & 9  & 5  & 5  & 2  & 1  & 1  & 6  & 9  & 9\\ 
	2  & 4  & 2  & 2  & 3  & 7  & 9  & 3  & 9  & 2  & 4  & 4  & 8  & 2  & 8  & 6  & 3  & 2\\ 
	6  & 2  & 5  & 5  & 1  & 2  & 2  & 2  & 2  & 6  & 1  & 4  & 9  & 5  & 7  & 1  & 2  & 6\\
   \hline
	LI & KB & FN & .C & BY & EB & SU & I. & BK & RN & KD & FD & BA & HQ & RP & X. & FT & AO\\
   \hline
   \end{tabular}
   \caption{Digrafid}
   \label{Digrafid-table-reference}
   \end{center}

>>>>

\item {\bf Shift cipher, for example Caesar cipher}\footnote{In CrypTool
   this method  can be find at three different places in the menu tree:\\
   - {\bf Crypt \textbackslash{} Symmetric (classic) \textbackslash{} 
      Caesar / ROT13}\\
   - {\bf Analysis \textbackslash{} Symmetric Encryption (classic)
     \textbackslash{} Ciphertext only \textbackslash{} Caesar} \\
   - {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     \textbackslash{} Caesar}. } 
   \cite{Singh2001}\index{Caesar}%
   : Plaintext alphabet and ciphertext alphabet are shifted against each 
   other by a determined number of letters. Using the Caesar cipher means 
   shifting letters about three positions.

@@

\item {\bf Nihilist Substitution}\footnote{An animation of this Nihilist method
   can be find in CrypTool at the menu item
     {\bf Indiv. Procedures \textbackslash{} Visualization of Algorithms 
     \textbackslash{} Nihilist}. }
   \cite{ACA2002}\index{Nihilist substitution}: 
   Insert the alphabet into a 5x5-matrix and replace 
   every letter of the message with the two corresponding digits. The 
   resulting numbers are written into a grille. For this purpose, a 
   keyword is chosen and placed above the columns of the grille. Its 
   letters are substituted by numbers, too. The ciphertext results from 
   adding the numbers of the plaintext and the numbers of the keyword.
   Numbers between 100 and 110 are transformed to numbers between 00 and
   10, so that each letter is represented by a two-digit number.

@@

   It is ostentatious, how often the numbers 1 and 2 appear,
   but this will be fixed with the following version.\\

@@

   \begin{center}
   \begin{tabular}{|c@{ }c@{ }c|c@{ }c@{ }c|c@{ }c@{ }c|c@{ }c@{ }c|c@{ }c@{ }c|c@{ }c@{ }c|}
   \hline		
	co & mb & in & in & gs & ub & st & it & ut & io & na & nd & tr & an & sp & os & it & io\\
   \hline
	3  & 4  & 9  & 9  & 7  & 3  & 1  & 9  & 3  & 9  & 5  & 5  & 2  & 1  & 1  & 6  & 9  & 9\\ 
	2  & 4  & 2  & 2  & 3  & 7  & 9  & 3  & 9  & 2  & 4  & 4  & 8  & 2  & 8  & 6  & 3  & 2\\ 
	6  & 2  & 5  & 5  & 1  & 2  & 2  & 2  & 2  & 6  & 1  & 4  & 9  & 5  & 7  & 1  & 2  & 6\\
   \hline
	LI & KB & FN & .C & BY & EB & SU & I. & BK & RN & KD & FD & BA & HQ & RP & X. & FT & AO\\
   \hline
   \end{tabular}
   \caption{Digrafid}
   \label{Digrafid-table-reference}
   \end{center}

### 2007-07-25 script/en/chapters/primes.tex
CT-scripts: Minor corrections -> should be final now for 1.4.10

The following table contains these Mersenne record primes\footnote{%
An up-to-date version can be found in the internet at
     \href{http://www.mersenne.org/history.htm}
  {\texttt{http://www.mersenne.org/history.htm}}.
}$^,$\footnote{%
Always, when a new record is published in the respective forums the same
and often ironic discussions start: Is there a deeper sense? Can this result
be applied for anything useful?
The answer is, that we don't know it yet. In fundamental research one
cannot see at once how it brings mankind forward.
}:

@@

% --------------------------------------------------------------------------
\vskip +10 pt
\hypertarget{generalizedFermatprimes}{}
\subsubsection[Generalized Fermat numbers $f(b,n) = b^{2^n} + 1$]
              {Generalized Fermat numbers\footnotemark~$f(b,n) = b^{2^n} + 1$}
    \footnotetext{%
      The base of this power is no longer restricted to 2 . \\
      Even more generic would be:  $f(b,c,n) = b^{c^n} \pm 1$
    }
\label{generalized-fermat}
\index{Fermat!number!generalized} 
    Generalized Fermat numbers are more numerous than Mersenne numbers of a
    equal size and many of them are waiting to be discovered to fill the
    big gaps between the Mersenne primes already found or still undiscovered.
    Progress in number theory made it possible that numbers, where the
    representation is not limited to the base 2, can be tested at almost
    the same speed than a Mersenne number.

@@

Such tables can easily be calculated with computer algebra systems.
Here a code sample for the CAS system Pari-GP\index{Pari-GP}:
\vspace{-2ex}
\begin{verbatim}
v2=[56, 64, 80, 90, 112, 128, 192, 256];
default(format,"e0.6");
for ( i=1,8, o=v2[i]; print(o, " --- ", 1.*2^o) );
\end{verbatim}

>>>>

The following table contains these Mersenne record primes\footnote{%
An up-to-date version can be found in the internet at
     \href{http://www.mersenne.org/history.htm}
  {\texttt{http://www.mersenne.org/history.htm}}.
}$^,$\footnote{%
Always, when a new record is published in the respective forums the same
and often ironic discussions start: Does this kind of research have a deeper
sense? Can this result be applied for anything useful?
The answer is, that we don't know it yet. In fundamental research one
cannot see at once whether and how it brings mankind forward.
%But this is the normal case with basic research, that one cannot see at once, 
}:

@@

% --------------------------------------------------------------------------
\vskip +10 pt
\hypertarget{generalizedFermatprimes}{}
\subsubsection[Generalized Fermat numbers $f(b,n) = b^{2^n} + 1$]
              {Generalized Fermat numbers\footnotemark~$f(b,n) = b^{2^n} + 1$}
    \footnotetext{%
      The base of this power is no longer restricted to 2.
      Even more generic would be:  $f(b,c,n) = b^{c^n} \pm 1$
    }
\label{generalized-fermat}
\index{Fermat!number!generalized} 
    Generalized Fermat numbers are more numerous than Mersenne numbers of a
    equal size and many of them are waiting to be discovered to fill the
    big gaps between the Mersenne primes already found or still undiscovered.
    Progress in number theory made it possible that numbers, where the
    representation is not limited to the base 2, can be tested at almost
    the same speed than a Mersenne number.

@@

Such tables can easily be calculated with computer algebra systems.
Here a code sample for the CAS system Pari-GP\index{Pari-GP}\index{CAS}:
\vspace{-2ex}
\begin{verbatim}
v2=[56, 64, 80, 90, 112, 128, 192, 256];
default(format,"e0.6");
for ( i=1,8, o=v2[i]; print(o, " --- ", 1.*2^o) );
\end{verbatim}
